Ripple Consensus Review

### 2015-04-25 paper.tex
Choosing the Unique Node List

\documentclass{article}
\usepackage{url}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
\begin{document}
\title{Ripple Consensus Review}
\author{Peter Todd}
\date{FIXME}
\maketitle

@@

Different nodes may chose to use different UNLs; if the UNLs do not
sufficiently overlap global consensus is not guaranteed because different UNL
``cliques'' can come to consensus independently of each other. For instance
Figure \ref{fig:unl-clique-disjoint} shows a simple example with two UNL
cliques, red and blue, that have forked because of insufficient connectivity,
resulting in the forked blockchain shown in Figure \ref{fig:unl-clique-disjoint-blockchain}.

@@

Regardless, it is certain that following a fork caused by disjoint UNL cliques
the losing side simply has to discard some or all history, as seen in Figure
\ref{fig:unl-clique-post-fork}, possibly resulting in double-spends and
associated financial losses.

@@

Similarly for a UNL of $n$ nodes it would
take $(4n + 1)/5$ Byzantine failures for an incorrect transaction to be
confirmed.

>>>>

\documentclass{article}
\usepackage{url}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{arrows,positioning}
\begin{document}
\title{Ripple Consensus Review}
\author{Peter Todd}
\date{FIXME}
\maketitle

@@

Different nodes may chose to use different UNLs; if the UNLs do not
sufficiently overlap global consensus is not guaranteed because different UNL
``cliques'' can come to consensus independently of each other. For instance
Figure \ref{fig:unl-clique-disjoint} shows a simple example with two UNL
cliques, red and blue, that have forked because of insufficient connectivity,
resulting in the forked blockchain shown in Figure
\ref{fig:unl-clique-disjoint-blockchain}. More complex failure modes have also
been identified, such as the ``less than majority evil'' failure mode
identified by Gregory Maxwell,\cite{gmaxwell-btctalk-ripple} and ``separate
supermajorities'' mode identified by Andrew Miller.\cite{amiller-rippletalk}

@@

Regardless, it is certain that following a fork caused by disjoint UNL cliques
the losing side simply has to discard some or all history (Figure
\ref{fig:unl-clique-post-fork}) possibly resulting in double-spends and
associated financial losses.

@@

\subsection{Choosing the Unique Node List}

How should Ripple node administrators pick their Unique Node List? Ripple Labs
provides little, if any, concrete advice beyond the ``starter UNL'' they
provide. Let's look at this problem from another angle: Why shouldn't a node
operator just use the default UNL?

Let's assume the node operator does not know what UNL other nodes in the
economic majority have chosen to use. They do however know the default UNL
provided by Ripple Labs, and they have a reasonable belief that other node
operators also know that UNL and may also be using it. The node operator's goal
is to limit losses; the best case is their node stays in consensus with other
nodes. Second best is they suffer a denial of service attack causing their node
to stop processing some or all transactions. Worst case is their node accepts
transactions that other nodes do not, making possible double-spend attacks.
Figure \ref{fig:unl-choice-plot} shows the outcomes of various possible
decisions by them and other node operators, ranging from $100\%$ default UNL,
$50\%$ default UNL, and $10\%$ default UNL.

In every circumstance node operators can reduce the risk of losses due to
consensus failures by removing non-default UNL entries, with the least risky
option being to simply stick with the default UNL of $100\%$ Ripple Labs
controlled nodes. This is particularly acute given the low profit margins of
many financial transactions: while a denial-of-service halts incoming revenue,
a double-spend attack can quickly drain working capital, quite possibly without
any hope of ever recovering the stolen funds.

A closely related issue is what incentive does a node operator have to publicly
perform validation services? Ripple currently has no compensation mechanism for
public validators, yet validation at minimum raises potential legal issues such
as lawsuits for negligence. Again, the least-risk option is to not publicly
validate the Ripple ledger.

\begin{table}
    \begin{tabular}{c|c|c|c|}
        ~ & \multicolumn{3}{c}{Majority UNL \%} \\
        \multirow{3}{*}{You} & $100\%$ & $50\%$ & $10\%$ \\ \hline
        $100\%$ & Consensus & DoS & DoS \\ \hline
        $50\%$ & DoS & DoS & D.S. \\ \hline
        $10\%$ & D.S & D.S & D.S. \\
    \end{tabular}
\end{table}
